Query Optimization for Dynamic Imputation

### 2017-05-14 rebuttal.tex

\documentclass{vldb}
\usepackage{xcolor}
\newcommand{\resp}[1]{{\color{blue}{#1}}}
\newcommand{\todo}[1]{{\color{red}{\textit{#1}}}}
\newcommand{\ProjectName}[0]{ImputeDB}

\begin{document}
\section{Introduction}
The authors would like to thank the reviewers for their helpful and insightful comments.
We have taken steps to address each comment and we believe that the paper is significantly improved.
In this document, we give a point-by-point response to the comments from each reviewer.

The reviewer's original comments are in black and our responses are in \resp{blue}.

In our draft, new text is highlighted in {\color{blue}blue} and changed text is highlighted in {\color{red}red}.

\section{Multiple reviewers}
\begin{itemize}
\item Redundant figures, excessive whitespace, formatting.

### 2017-05-14 rebuttal.tex

\section{Multiple reviewers}
\begin{itemize}
\item Redundant figures, excessive whitespace, formatting.

  \resp{We've compressed many of the figures and removed whitespace to make room for additional material.}
\end{itemize}

  \section{Reviewer 1}
  \begin{itemize}
  \item W2. The paper completely obscures the ML technique being used: (1) It would be nice to have some discussion of the complete API for implementing an ML technique, as well as how one might implement a cost-estimator for one or two examples. (2) It would be helpful to see the impact of a few different approaches (e.g., Tree-based, Interpolation, Regression, FD-based) on performance and accuracy. (3) Given that the ML time estimator is being treated as a black box, it would be useful to see how sensitive the plan selector is to inaccurate estimates.

  \resp{
    We extended our evaluation to include two commonly used imputation methods: mean value imputation and hot-deck imputation.
    These methods are in addition to the original imputation method (CART), which is based on regression trees.
    The regression tree model incorporates correlations between attributes, so should be able to take advantage of any functional dependencies to make better predictions.
    Interpolation based imputation is useful for data where the ordering of tuples is significant, like time series, but is not well suited for survey data where the ordering of tuples is arbitrary.

    We also added details on how to pick the cost functions that define the API for impute operators.
    We provide details on how these functions are defined for all the methods used in our evaluation.

    Our optimizer (like any query optimizer) is sensitive to the quality of its estimates.
    We believe that the cost functions we have chosen lead to good results overall, but there may be cases where our estimates lead to bad outcomes.
    It is an interesting and difficult problem to accurately estimate the quality of a regression model before running the model.
    Any work in this area could be used immediately to improve our algorithm.
  }

### 2017-05-14 rebuttal.tex

\documentclass{vldb}
\usepackage{xcolor}
\newcommand{\resp}[1]{{\color{blue}{#1}}}
\newcommand{\todo}[1]{{\color{red}{\textit{#1}}}}
\newcommand{\ProjectName}[0]{ImputeDB}

\begin{document}
\section{Introduction}
The authors would like to thank the reviewers for their helpful and insightful comments.
We have taken steps to address each comment and we believe that the paper is significantly improved.
In this document, we give a point-by-point response to the comments from each reviewer.

In the rebuttal, the reviewer's original comments are in black and our responses are in \resp{blue}.

@@

In our draft, new text is highlighted in {\color{blue}blue} and changed text is highlighted in {\color{red}red}.

\section{Overall changes}
\begin{itemize}
\item Redundant figures, excessive whitespace, formatting.

  \resp{Several reviewers commented on the amount of whitespace, so we've compressed many of the figures and removed whitespace to make room for additional material.}

\item \resp{
    We have improved our system since the paper was submitted by adding imputation for categorical variables.
    This has a minor impact on the experimental results.
    It improves the ability of the regression tree model to perform imputation by making these variables available to the model.
  }
\end{itemize}

### 2017-05-14 paper/algorithms/histogram-transform.tex
paper: removed histogram helpers and tried to rework into simpler form
paper: fixed general typos, added clearer todos
paper: revise histogram algorithm
paper: pass over cardinality estimation section
paper: highlight major changes & additions

%\begin{algorithm}
  \begin{algorithmic}
    \Require{$\mathbb{H}$ is a map from attribute names to histograms, $op$ is an operator node in a logical query plan}
    \Ensure{Returns an updated map $\mathbb{H'}$}

    \Function{UpdateHistograms}{$\mathbb{H}, op$}
    	\If{$op \not\in \{\delta, \mu, \sigma, \bowtie \}$}
		\Return{$\mathbb{H}$}
	\EndIf
	\State $\mathbb{H'} \gets \Call{Copy}{\mathbb{H}}$
    	\If{op = $\delta_C$}
		\For{$c \in C$}
			\State $\mathbb{H'}[c][null] \gets 0$
		\EndFor
	\ElsIf{op = $\mu_C$}
		\For{$c \in C$}
			\State \Call{Add}{$\mathbb{H'}[c], \mathbb{H'}[c][null]$}
		\EndFor
	\ElsIf{op = $\sigma$}
		\For{$c \in dom(\mathbb{H'})$}
			\State \Call{ScaleBy}{$\mathbb{H'}[c],  \Call{Selectivity}{\sigma}$} 
		\EndFor
	\ElsIf{op = $\bowtie_\phi$}
		\For{$c \in dom(\mathbb{H'})$}
			\State \Call{ScaleTo}{$\mathbb{H'}[c], \Call{Cardinality}{\bowtie_{\phi}}$}
		\EndFor
	\EndIf
	\Return{$\mathbb{H'}$}
    \EndFunction
  \end{algorithmic}
  \caption{An algorithm for in-plan histogram updates}
  \label{algo:histogram-transformation}
%\end{algorithm}

>>>>

\newcommand{\InlineIte}[3]{#2\ \textbf{if}\ #1\ \textbf{else}\ #3}
\renewcommand{\algorithmicindent}{1em}

\review{
\begin{algorithmic}[1]
    \Require{
    \Statex
    \begin{itemize}[leftmargin=*,noitemsep]
    \item $H$: A map from attribute names to histograms.
    \item $op$: A relational operator.
    \end{itemize}}

  \Function{UpdateHistograms}{$H,\ op$}
  \If{op = $\delta_C$}\label{lst:drop-start}
  \State Set null count to zero for $H[c]$ where $c \in C$.
  \State Scale $H[c]$ s.t. $\Call{Card}{H[c]} = \min_d \Call{Card}{H[d]}$, $\forall c$.\label{lst:drop-end}
  \ElsIf{op = $\mu_C$}\label{lst:impute-start}
  \State Scale $H[c]$ s.t. $\Call{Card}{H[c]}$ is unchanged and null
  \State count is zero, $\forall c$.\label{lst:impute-end}
  \ElsIf{op = $\sigma_\phi$}\label{lst:std-start}
  \State Scale $H[c]$ by $\Call{Selectivity}{\phi, H}$, $\forall c$.
  \ElsIf{op = $\bowtie_\phi$}
  \State Scale $H[c]$ s.t. $\Call{Card}{H[c]} = \Call{Card}{\bowtie_\phi}$, $\forall c$.
  \EndIf\label{lst:std-end}
  \State \Return $H$
  \EndFunction
\end{algorithmic}
}
\caption{\review{An algorithm for in-plan histogram updates}}
\label{algo:histogram-transformation}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:

### 2017-05-14 paper/examples.tex
paper: highlight major changes & additions

An alternative strategy involves starting with $\alpha=1$ and iterating to lower values, as long as execution
of the query completes within some user-defined time bound. \ProjectName{}'s planning
algorithm can expose the final set of plans that tradeoff execution time and imputation quality, which allows
the user to adjust their choice of $\alpha$ to uniquely select amongst these plans.

>>>>

\review{
An alternative strategy involves starting with $\alpha=1$ and iterating to lower values, as long as execution
of the query completes within some user-defined time bound. \ProjectName{}'s planning
algorithm can expose the final set of plans that tradeoff execution time and imputation quality, which allows
the user to adjust their choice of $\alpha$ to uniquely select amongst these plans.
}

### 2017-05-14 paper/experiments.tex
paper: highlight major changes & additions

\Cref{fig:pareto-frontiers} shows the final Pareto frontiers produced for \Cref{q1} and \Cref{q6}. These are the resulting set of plans after pruning dominated plans throughout our algorithm.
and encapsulate the possible tradeoffs that the user can make in choosing a final plan for execution. The frontier is available at the end of the planning stage and
can be exposed to the user to help in guiding their workflow.

@@

\begin{figure}
\centering
\includegraphics[scale=0.4]{figures/pareto_frontiers_plot.png}
\caption{\ProjectName{}'s final Pareto frontier for \Cref{q1} and \Cref{q6}. 
For clarity, we have removed from the plot plans that provide nearly indistinguishable tradeoffs in either dimension.}
\label{fig:pareto-frontiers}
\end{figure}

@@

In contrast, \ProjectName{} executes the following query:
\begin{lstlisting}[breaklines]
SELECT AVG(c0) FROM acs;
\end{lstlisting}
over the ACS dataset (also using chained-equation decision trees imputation) in \acsimputedbzeroresult{} when optimizing for quality and \acsimputedboneresult{} when optimizing for runtime.
The performance difference comes from the fact that \ProjectName{} only needs to perform imputation on the single column required by the query.
Adding selection predicates would further reduce the query runtime by reducing the amount of required imputation.
An analyst could do the same, but tracking the required imputations would get increasingly complicated as the queries became more complex.
This highlights the benefit of using our system for early data exploration.

@@

However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
We explored the limits of our planning algorithm by constructing a series of queries with increasing number of joins over the CDC tables.
Planning times (~\Cref{fig:planning-times}) for queries involving 1--5 joins---a practical value for  real world exploratory queries---are less than 1 second.
We extended \ProjectName{} to support an approximation of Pareto sets to improve planning times for queries with 6 or more joins. Approximate Pareto sets
displayed a reduction in planning time linear with respect to the reduction in size of the final frontier. For queries with 6--8 joins, approximate sets achieved a reduction
in planning time between 14\% and 23\%, on average. Users can increase the level of approximation to reduce planning times further.
%\todobox{Jose: I'm not a fan of this line, I would remove}{Furthermore, at that point, it may make more sense to specify a coherent model over the entire base table instead.}

>>>>

\review{
\Cref{fig:pareto-frontiers} shows the final Pareto frontiers produced for \Cref{q1} and \Cref{q6}. These are the resulting set of plans after pruning dominated plans throughout our algorithm.
and encapsulate the possible tradeoffs that the user can make in choosing a final plan for execution. The frontier is available at the end of the planning stage and
can be exposed to the user to help in guiding their workflow.
}

@@

\begin{figure}
\centering
\includegraphics[scale=0.4]{figures/pareto_frontiers_plot.png}
\caption{\review{
    \ProjectName{}'s final Pareto frontier for \Cref{q1} and \Cref{q6}. 
    For clarity, we have removed from the plot plans that provide nearly indistinguishable tradeoffs in either dimension.
}}
\label{fig:pareto-frontiers}
\end{figure}

@@

In contrast, \ProjectName{} executes the query \lstinline{SELECT AVG(c0) FROM acs} over the ACS dataset (also using chained-equation decision trees imputation) in \acsimputedbzeroresult{} when optimizing for quality and \acsimputedboneresult{} when optimizing for runtime.
The performance difference comes from the fact that \ProjectName{} only needs to perform imputation on the single column required by the query.
Adding selection predicates would further reduce the query runtime by reducing the amount of required imputation.
An analyst could do the same, but tracking the required imputations would get increasingly complicated as the queries became more complex.
This highlights the benefit of using our system for early data exploration.

@@

\review{
However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
We explored the limits of our planning algorithm by constructing a series of queries with increasing number of joins over the CDC tables.
Planning times (~\Cref{fig:planning-times}) for queries involving 1--5 joins---a practical value for  real world exploratory queries---are less than 1 second.
We extended \ProjectName{} to support an approximation of Pareto sets to improve planning times for queries with 6 or more joins. Approximate Pareto sets
displayed a reduction in planning time linear with respect to the reduction in size of the final frontier. For queries with 6--8 joins, approximate sets achieved a reduction
in planning time between 14\% and 23\%, on average. Users can increase the level of approximation to reduce planning times further.
}
%\todobox{Jose: I'm not a fan of this line, I would remove}{Furthermore, at that point, it may make more sense to specify a coherent model over the entire base table instead.}
